Quality Metrics for SKOS Vocabularies

### 2012-03-15 background.tex
polished background section

\section{Background and Methodology}

@@

SKOS is a language for defining vocabularies in the Web of Data and therefore based on the Open World Assumption. Established quality notions from closed-world systems, such as referential integrity or schema validation, don't hold anymore, because the available information may be incomplete and facts that are not explicitly stated cannot be determined as true or false. While trust and provenance models for Web data are being developed~\cite{Omitola2011,Hartig2009}, content-based and hand-crafted heuristics are currently used to evaluate quality in Linked Data sets~\cite{Heath2011}. In the course of the LATC project, Gueret et al. propose an infrastructure that utilizes crowd-sourcing technology to semi-automatically create links between datasets of the LOD cloud. To ensure the quality of the resulting network, human judgement is supported by calculating a report of locally approximated network measures. \todo{BH}{ich denke antoine hat das paper gemeint: http://www.mendeley.com/download/public/18928/4084690763/56bdde328dfbba2cb8594b57152c0e132ef00a40/dl.pdf . Leider fehlen da etliche Details um sich ein konkretes Bild machen zu können}

@@

Interoperability between applications has been the major motivation for specifying in total six \textbf{SKOS integrity conditions}~\cite{Miles2005}. An integrity condition is a statement that defines under which circumstances data are consistent with the SKOS data model. An example condition is ``A resource has no more than one value of skos:prefLabel per language tag''. Tools can check whether these integrity conditions are met for given data. Two of the six integrity conditions are defined formally in the OWL representation of SKOS and can therefore be verified by any OWL reasoner. Poolparty and SKOSify implement all these integrity checks plus \todo{CM}{summarize how existing tools implement these conditions}

% Argue why we need higher-level criteria

\todo{CM}{argue why quality criteria beyond integrity conditions are needed; provide some good arguments and refer to the literature. The focus of this paragraph should be on quality in the controlled vocabulary domain...}

% Methodology

Our goal was to identify quality criteria that go beyond SKOS integrity conditions. These criteria do not primarily address interoperability on the data model level but usability and interoperability in the application areas mentioned before. We used the following sources for our investigation:

\begin{itemize}
    \item Existing literature
    \item Mailing list discussions
    \item \ldots \todo{cm}{think about and finish this paragraph and lists.}
\end{itemize}

After we identified an initial set of criteria, we formalized them into computable metrics and applied them on existing SKOS vocabularies to learn about the real-world occurrences of these criteria. During that process we also learned about additional quality issues in real world SKOS data, refined already identified criteria and included additional ones in a SKOS quality criteria catalogue, which we published online\footnote{\url{https://github.com/cmader/qSKOS/wiki/Quality-Criteria-for-SKOS-Vocabularies}}.

>>>>

\section{Background}

@@

SKOS is a language for defining vocabularies in the Web of Data and therefore based on the Open World Assumption. Established quality notions from closed-world systems, such as referential integrity or schema validation, don't hold anymore, because the available information may be incomplete and facts that are not explicitly stated cannot be determined as true or false. While trust and provenance models for Web data are being developed~\cite{Omitola2011,Hartig2009}, content-based and hand-crafted heuristics are currently used to evaluate quality in Linked Data sets~\cite{Heath2011}. Gueret et al. [\todo{CM}{find the right reference in google scholar}], for instance, calculate network measures to support human judgement on the quality of Linked Data \todo{CM}{?} graphs.

%In the course of the LATC project, Gueret et al. propose an infrastructure that utilizes crowd-sourcing technology to semi-automatically create links between datasets of the LOD cloud. To ensure the quality of the resulting network, human judgement is supported by calculating a report of locally approximated network measures. \todo{BH}{ich denke antoine hat das paper gemeint: http://www.mendeley.com/download/public/18928/4084690763/56bdde328dfbba2cb8594b57152c0e132ef00a40/dl.pdf . Leider fehlen da etliche Details um sich ein konkretes Bild machen zu können}

@@

Interoperability between applications has been the major motivation for specifying in total six \textbf{SKOS integrity conditions}~\cite{Miles2005}, each of which is a statement that defines under which circumstances data are consistent with the SKOS data model. An example condition is ``a resource has no more than one value of skos:prefLabel per language tag''. Tools that can check whether these integrity conditions are met for given data are already available: two of the six integrity conditions are defined formally in the OWL representation of SKOS and can therefore be validated by any OWL reasoner \todo{CM}{true?}. For validating a SKOS vocabulary against the integrity conditions, one can use tools such as the online \textbf{PoolParty Thesaurus Consistency Checker}, or the \textbf{Skosify} command line utility, which can validate SKOS vocabularies and also correct some detected quality problems.

% Argue why we need criteria that go beyond low-level integrity constraints and why we need to formalize them

Typical application areas of controlled vocabularies are as classification, indexing, autocompletion, query reformulation/expansion, or serving as a glossary. As we discussed in detail in our earlier work [\ref{CM}{iSem paper?}], these application areas have specific requirements with respect to vocabulary features, such as structure, availability, and documentation of conceptual resources. A vocabulary that doesn't fulfill these requirements can have a negative effect on an application's output and would therefore be judged by the application designers and potentially also by users as being a low-quality vocabulary. Missing relations between conceptual resources, for instance, can reduce retrieval recall when a vocabulary is used for query expansion. Missing entry points to concepts in a vocabulary can impede orientation for human users. Inconsistent relations can hinder queries across distinct datasets\todo{cm}{this example is unclear}, to name some examples.

% Argue why we need to formalize them -> because otherwise they will be implemented inconsistently across applications
Thus, in order to help vocabulary and application designers in judging the quality of SKOS vocabularies, we need higher-level quality criteria that go beyond low-level integrity conditions, which define data model consistency. Such quality criteria are already well-established and have extensively been discussed in standardized guidelines [\todo{CM}{ISO refs}], manuals [\todo{CM}{hedden book, skosify ref 3}], and scholarly articles [\todo{cm}{e.g., skosify refs 16/17}].
% TODO: bitte hier einfach nur die referenzen auf die wir uns beziehen zusammenfassen und kategorisieren. Schau dir dazu auch nochmals die SKOSify referenzen und die aus eurem iSemantics paper an.
If we want to apply these criteria on SKOS vocabularies, we first need to identify them and determine how to evaluate them given the syntax, structure and semantics of SKOS. Since we want to provide tool support for quality checking, we must formalize these criteria into computable metrics, in order to ensure consistent implementation and interpretation across tools.

% I think for the big picture it is not necessary to go into such details. We can do this later when introducting the metrics

% The PoolParty Thesaurus Consistency Checker implements all these conditions and introduces the following new checks: \textbf{URI validation} checks for invalid characters like, e.g. whitespaces in the URI string. \textbf{Definition of labels} is tested for conceptual resources and concept schemes. Furthermore, SKOS labels and textual content is checked for \textbf{missing language tags} and the notion of \textbf{loose concepts} is defined: it encompasses conceptual resources that are no top concepts and have no broader concept.
% 
% The SKOSify tool implements three of the above mentioned SKOS integrity conditions as well as the \textbf{missing language tags} and \textbf{loose concepts} measures introduced by the PoolParty checker. It introduces \textbf{hierarchy cycle detection} and detection of \textbf{extra whitespace} for SKOS label or documentation properties.
